---
title: "General Model - Household Living Budget - HLB"
author: "cm"
date: "09/13/2023" #update from the previous code
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Start fresh
```{r}
rm(list=ls())
```

# Libraries
```{r}
#install.packages("ipumsr")
library(ipumsr)
#install.packages("tidyverse")
#library(tidyverse)
library(survival)
library(ggfortify)
library(dplyr)
library(readxl)
library(readr)
library(reshape2)
library(tidyr)
```

#  ######################## 1 DETAILS: Specify the fips code for the State and county ########################
```{r}
state_name <- "va"

county_fips <- 51059  #Fairfax County, VA 

county_name <- "Fairfax County"

counties_shared_puma <- c(51059, 51610, 51600) #PUMAS include Fairfax County, Fairfax city and Falls Church

name_files <- "fairfax"
```

#Root folder of the repo RFR
```{r}
#directory of the file
dirname(rstudioapi::getSourceEditorContext()$path)
#dirname finds the root folder of the current document (fairfax_hlb.Rmd)
root_repo <- dirname(dirname(dirname(rstudioapi::getSourceEditorContext()$path)))
```


## Data PUMS
```{r}
#url
url_xml <- "https://raw.githubusercontent.com/uva-bi-sdad/household_living_budget/main/documents/products/derived_variables/fairfax_va_acs_pums_derived_variables_2021.csv"

#read csv from url
dat_hlb <- read_csv(url_xml)

```

# Crosswalks Census Tracts - PUMA
```{r}
#source https://www.census.gov/programs-surveys/geography/guidance/geo-areas/pumas.html
#source tallies https://www2.census.gov/geo/docs/maps-data/data/geo_tallies2020/tallies_by_state/Virginia_51.txt 

# puma_ct <- read_delim("~/git/household_living_budget/data/demographic/crosswalks/2010_Census_Tract_to_2010_PUMA.txt", delim = ",")
# puma_ct_20 <- read_delim("~/git/household_living_budget/data/demographic/crosswalks/2020_Census_Tract_to_2020_PUMA.txt", delim = ",")

# puma_ct: PUMAS 2010
filepath <- paste0(root_repo, "/data/demographic/crosswalks/2010_Census_Tract_to_2010_PUMA.txt")
puma_ct <- read_delim(filepath, delim = "," )
# puma_ct_20: PUMAS 2020
filepath <- paste0(root_repo, "/data/demographic/crosswalks/2020_Census_Tract_to_2020_PUMA.txt")
puma_ct_20 <- read_delim(filepath, delim = ",")
  
# crosswalk for Fairfax County only
puma_ct_fairfax <- puma_ct %>% filter(STATEFP=='51', COUNTYFP== '059')

#puma_ct_fairfax_20 <- puma_ct_20 %>% filter(STATEFP=='51', COUNTYFP== '059')

puma_ct_fairfax_20_county <- puma_ct_20 %>% filter(STATEFP=='51', COUNTYFP== '059')
puma_ct_fairfax_20_city <- puma_ct_20 %>% filter(STATEFP=='51', COUNTYFP== '610')
puma_ct_fairfax_20_fallschurch <- puma_ct_20 %>% filter(STATEFP=='51', COUNTYFP== '600')

puma_ct_fairfax_20 <- rbind(puma_ct_fairfax_20_county, puma_ct_fairfax_20_city, puma_ct_fairfax_20_fallschurch)

#puma_ct_fairfax_2 <- puma_ct_fairfax %>% head(2)
puma_ct_fairfax$fips <- paste0(puma_ct_fairfax$STATEFP, puma_ct_fairfax$COUNTYFP, puma_ct_fairfax$TRACTCE )
puma_ct_fairfax_20$fips <- paste0(puma_ct_fairfax_20$STATEFP, puma_ct_fairfax_20$COUNTYFP, puma_ct_fairfax_20$TRACTCE )

#comparison
compare_puma_ct <- puma_ct_fairfax_20 %>% left_join(puma_ct_fairfax, by='fips')
table(compare_puma_ct$PUMA5CE.x, compare_puma_ct$PUMA5CE.y)
equivalence <-  compare_puma_ct %>% group_by(PUMA5CE.x) %>% summarise(equiv=first(PUMA5CE.y))

#rewrite categories of puma_ct_fairfax_20 (274 tracts) to match puma_ct_fairfax (258 tracts)
names(puma_ct_fairfax_20)[4] <- 'PUMA5CE_20'

puma_ct_fairfax_20 <- puma_ct_fairfax_20 %>% mutate(PUMA5CE= case_when(PUMA5CE_20 == '05901' ~ '59309',
                                                                       PUMA5CE_20 == '05902' ~ '59307',
                                                                       PUMA5CE_20 == '05904' ~ '59305',
                                                                       PUMA5CE_20 == '05905' ~ '59304',
                                                                       PUMA5CE_20 == '05906' ~ '59306',
                                                                       PUMA5CE_20 == '05907' ~ '59301',
                                                                       PUMA5CE_20 == '05908' ~ '59302',
                                                                       PUMA5CE_20 == '60001' ~ '59303',
                                                                       PUMA5CE_20 == '61001' ~ '59308'
                                                      ) )

```


#  ######################## 2 Household Living Budget Components - HLB  ########################

# 2.1 Exercise at the household level

# Food Cost 
```{r}
#source:
#https://www.fns.usda.gov/cnpp/usda-food-plans-cost-food-reports-monthly-reports
#https://fns-prod.azureedge.us/sites/default/files/media/file/CostofFoodAug2022LowModLib.pdf
#adjustmen Feeding America
#https://map.feedingamerica.org/county/2020/overall/virginia

#creat path to the file
filepath <- paste0(root_repo, "/data/household_living_budget/food/va_ct_feeding_america_map_meal_gap_2022SEP.xlsx")
#read file
food_cost_adj <- read_excel(filepath)

#data USDA USA
filepath <- paste0(root_repo, "/data/household_living_budget/food/national_usda_meal_plans_2022SEP.csv")
usda_meal_plan_sep2022 <- read_csv(filepath, skip = 1)


#data USDA categories and cost
usda_low_cost_meal <- usda_meal_plan_sep2022 %>% select(group=`...2`, monthly_cost=`Low-cost...6`) %>% filter( !is.na(monthly_cost)) %>% filter(monthly_cost!= 'plan')

#remove $
usda_low_cost_meal$monthly_cost <- gsub( '\\$' ,'', usda_low_cost_meal$monthly_cost )

#vector with costs
food_cost_general <-  data.frame(
no_adult=  mean( c(as.numeric( usda_low_cost_meal %>% filter(group=='19-50 years') %>% pull(monthly_cost) ) ) ),
no_infant= as.numeric( usda_low_cost_meal %>% filter(group=='1 year') %>% pull(monthly_cost) ) ,
no_toddler=  mean( c(as.numeric( usda_low_cost_meal %>% filter(group=='2-3 years') %>% pull(monthly_cost) ) ) ),
no_preschooler= mean(as.numeric(usda_low_cost_meal %>% filter(group=='4-5 years') %>% pull(monthly_cost)) ),
no_schooler= mean(c( as.numeric(usda_low_cost_meal %>% filter(group=='6-8 years') %>% pull(monthly_cost)) ,
                     as.numeric(usda_low_cost_meal %>% filter(group=='9-11 years') %>% pull(monthly_cost)) )),
no_teenager=  mean(c( as.numeric(usda_low_cost_meal %>% filter(group=='12-13 years') %>% pull(monthly_cost)) ,
                     as.numeric(usda_low_cost_meal %>% filter(group=='14-18 years') %>% pull(monthly_cost)) )) )


#ADJUSTMENTS
#AVERAGE MEAL COST IN THE UNITED STATES: $3.25. Source: Feeding America for 2020
meal_avg_cost_us <- 3.25
# Fairfax Average Meal Cost: $4.23
#meal_avg_cost_fairfax_va <- 4.23
meal_avg_cost_fairfax_va <- food_cost_adj %>% filter(FIPS==county_fips) %>% pull(`Cost Per Meal (1 Year)`)

#adjustment
adj_food_cost <- meal_avg_cost_fairfax_va /meal_avg_cost_us
#food inflation - BLS to adjust meal cost from 2020
inflation_usa_food_2021 <- 0.065
inflation_usa_food_2022 <- 0.118
inflation_msa_mdv_food_2021 <- 0.057
inflation_msa_mdv_food_2022 <- 0.087

#adjustment + inflation_food adjustment
adj_food_cost_inf <- meal_avg_cost_fairfax_va/meal_avg_cost_us*(1+inflation_msa_mdv_food_2021)*(1+inflation_msa_mdv_food_2022)/((1+inflation_usa_food_2021)*(1+inflation_usa_food_2022))

#food cost adjusted by geography (county) and by inflation
food_cost_general_adj_infl <- adj_food_cost_inf*food_cost_general


#vector of per person cost - food
vector_food_cost <- data.frame(cost_pc = t(food_cost_general_adj_infl))
vector_food_cost <- tibble::rownames_to_column(vector_food_cost, "person")

#food cost long format
food_cost_long <- dat_hlb  %>% select(puma, serialno, hh_income_cat, hh_size_recode, no_adult, no_preschooler, no_toddler, no_schooler, no_teenager, no_infant ) %>% arrange(serialno) %>% pivot_longer(cols=c('no_adult', 'no_preschooler', 'no_toddler', 'no_schooler', 'no_teenager', 'no_infant'),
                    names_to='person',
                    values_to='num_person') %>% left_join(vector_food_cost, by = "person") %>% mutate(adj_econ_scales = case_when( 
                      hh_size_recode == 1 ~ 1 + 0.2, 
                      hh_size_recode == 2 ~ 1 + 0.1,
                      hh_size_recode == 3 ~ 1 + 0.05,
                      hh_size_recode == 4 ~ 1 + 0,
                      hh_size_recode == 5 ~ 1 - 0.05)
                      ) %>% mutate(food_cost_month = num_person*cost_pc*adj_econ_scales)

#food cost
food_hlb <- food_cost_long %>% group_by(serialno) %>% summarise(food_cost_month = sum(food_cost_month))

```


# Childcare Cost
```{r}
#data -  Dept of Labor
filepath <- paste0(root_repo, "/data/household_living_budget/childcare/dmv_dol_childcare_costs_2022.xlsx")
childcare_dol <- read_excel(filepath)

#data county
childcare_dol_county <- childcare_dol %>% filter(year==2022) %>% filter(FIPS == county_fips) %>%
  select("name_county", "FIPS", "year", "infant_family", "toddler_family", "preschooler_family", "schooler_family" )


#weekly cost
cc_cost_indiv_min <- c(
adult= 0, 
infant= childcare_dol_county %>% pull(infant_family),
toddler= childcare_dol_county %>% pull(toddler_family),
preschooler= childcare_dol_county %>% pull(preschooler_family),
schooler= childcare_dol_county %>% pull(schooler_family),
teenager=0) 


# cost of child care per representative family - min - per month
childcare_hlb <-  dat_hlb %>% mutate(childcare_cost_month= 
                                                         as.numeric(cc_cost_indiv_min['adult'])*no_adult*4  +
                                                         as.numeric(cc_cost_indiv_min['teenager'])*no_teenager*4 +
                                                         as.numeric(cc_cost_indiv_min['schooler'])*no_schooler*4 +
                                                         as.numeric(cc_cost_indiv_min['preschooler'])*no_preschooler*4 +
                                                         as.numeric(cc_cost_indiv_min['toddler'])*no_toddler*4 +
                                                         as.numeric(cc_cost_indiv_min['infant'])*no_infant*4
                                                         ) %>% select(serialno, childcare_cost_month )

```

#Merge with df_hlb
```{r}

#join food cost
df_hlb <-  dat_hlb %>% left_join(food_hlb %>% select(serialno, food_cost_month) , by = 'serialno')

#childcare
df_hlb <- df_hlb %>% left_join( as.data.frame(childcare_hlb) %>% select(serialno, childcare_cost_month) , by = 'serialno') %>% select(-c("...1"))

#preliminary
#write.csv(df_hlb, "~/git/household_living_budget/documents/products/derived_variables/fairfax_va_puma_hlb_temporary_2021.csv", row.names = FALSE )

```


#Fill in information matrix tract-income-combinations_hh (tic)
```{r}
#df_0_actual vs (tract, puma)

#geoid AND puma
geoid_puma <- puma_ct_fairfax_20 %>% filter(COUNTYFP=="059") %>% select(geoid=fips, puma=PUMA5CE)
df_hlb$puma <- as.character(df_hlb$puma)

#joing geoid_puma with df_hlb: we are saying that the structure of every puma replicates for every census tract
# df_hlb_tic <- geoid_puma %>% left_join(df_hlb %>% select(puma, serialno, income_cat, hhincome, hh_recode, hh_compos, adult, num_hh=pop, food_cost_month, healthcare_cost_month, childcare_cost_month), 
#                                         by = c("puma"))
#changed
df_hlb_tic <- geoid_puma %>% left_join(df_hlb %>% select(puma, serialno, hh_income_cat, hh_income, hh_size_recode, hh_comb, no_adult, hh_wt, food_cost_month, childcare_cost_month), 
                                        by = c("puma"))


```


#2.2 Exercise at the tract level: housing and transportation
# Housing Cost 
```{r}
#data fro HUD
filepath <- paste0(root_repo, "/data/household_living_budget/housing/va_tr_hud_2022_housing_cost.xlsx")
housing_cost_va_updated <- read_excel(filepath) %>% select(-c(`...1`))

housing_cost_va_updated <- housing_cost_va_updated %>% mutate(fips=substr(geoid,1,5)) %>% filter(fips %in% counties_shared_puma )

housing_cost_va_updated_wide <- spread(housing_cost_va_updated, key = measure, value = value)

housing_cost_county <- housing_cost_va_updated_wide %>% mutate(
  `1` =  monthly_rent_0br,
  `2` =  monthly_rent_1br,
  `3` =  monthly_rent_2br,
  `4` =  monthly_rent_3br,
  `5` =  monthly_rent_4br
  ) %>% select( -c(monthly_rent_0br,monthly_rent_1br,monthly_rent_2br,monthly_rent_3br,monthly_rent_4br, measure_type, region_type, region_name, year, fips) )


#housing cost in long format
housing_cost_long <- housing_cost_county %>% pivot_longer(cols=c(`1` ,  `2`  , `3` ,  `4` ,  `5`),
                    names_to='hh_size',
                    values_to='housing_cost_month')

```


# Cost of Transportation
```{r}
#data va 2022 updated
filepath <- paste0( root_repo, "/data/household_living_budget/transportation/va_tr_cnt_transportation_cost_2022.xlsx")
transp_cost_va <- read_excel(filepath)

####
#create fips
transp_cost_va$fips <- substr(transp_cost_va$tract, 1, 5)

#data county per tract
transp_cost_va$tract <- as.character(transp_cost_va$tract)

transp_cost_county_tract <- transp_cost_va %>% filter(fips %in% counties_shared_puma) %>% select(tract, Transport_1) %>% 
     left_join(puma_ct_fairfax_20 %>% select(puma=PUMA5CE, fips) , by=c('tract'='fips'))

```

#Broadband cost 
```{r}
#data
filepath <- paste0(root_repo, "/data/household_living_budget/broadband/dcmdva_hdcttrbg_2021_broadband_now_internet_package_price.csv.xz")
dat_broadband <- read.csv(filepath)

#county id
dat_broadband$county <- substr(dat_broadband$geoid,1,5)

library(dplyr)
#use min price 25
bband_county_tr <- dat_broadband %>% filter(county %in% counties_shared_puma) %>% filter(region_type =='tract') %>% filter(measure=="min_price_25")

price_increase_wireless_service <- 0  #0.013
bband_tract <- bband_county_tr %>% mutate(broadband_cost_month = value*(1+price_increase_wireless_service)) %>% select(geoid, broadband_cost_month)

```


#Merge with df_hlb_tic
```{r}
#housing
df_hlb_tic$hh_size_recode <-  as.character(df_hlb_tic$hh_size_recode)
df_hlb_tic <- df_hlb_tic %>% left_join(housing_cost_long, by = c("geoid", "hh_size_recode"= "hh_size") )

#transportation
df_hlb_tic <- df_hlb_tic %>% left_join(transp_cost_county_tract %>% select(tract, Transport_1), by = c("geoid"="tract") ) %>% mutate(transp_cost_month = no_adult*Transport_1) 
df_hlb_tic <- df_hlb_tic %>% select(-c(Transport_1))

#NAs - average by hh_size_recode
complement_trasp <- df_hlb_tic %>% select(hh_size_recode, transp_cost_month) %>% group_by(hh_size_recode) %>% summarise(mean_transp = mean(transp_cost_month, na.rm = TRUE))

#complete NAs by hh_size_recode
df_hlb_tic <- df_hlb_tic %>% mutate(transp_cost_month = case_when( !is.na(transp_cost_month) ~ transp_cost_month, 
                                                     is.na(transp_cost_month) & hh_size_recode==1 ~ complement_trasp %>% filter(hh_size_recode==1) %>% select(mean_transp) %>% pull(),
                                                     is.na(transp_cost_month) & hh_size_recode==2 ~ complement_trasp %>% filter(hh_size_recode==2) %>% select(mean_transp) %>% pull(),
                                                     is.na(transp_cost_month) & hh_size_recode==3 ~ complement_trasp %>% filter(hh_size_recode==3) %>% select(mean_transp) %>% pull(),
                                                     is.na(transp_cost_month) & hh_size_recode==4 ~ complement_trasp %>% filter(hh_size_recode==4) %>% select(mean_transp) %>% pull(),
                                                     is.na(transp_cost_month) & hh_size_recode==5 ~ complement_trasp %>% filter(hh_size_recode==5) %>% select(mean_transp) %>% pull()
                                                     ))


#broadband - internet
df_hlb_tic <- df_hlb_tic %>% left_join(bband_tract, by = c("geoid") )

#NAs - average by hh_size_recode
complement_bband <- df_hlb_tic %>% select(hh_size_recode, broadband_cost_month) %>% group_by(hh_size_recode) %>% summarise(mean_bband = mean(broadband_cost_month, na.rm = TRUE))

#complete NAs by hh_size_recode
df_hlb_tic <- df_hlb_tic %>% mutate(broadband_cost_month = case_when( !is.na(broadband_cost_month) ~ broadband_cost_month, 
                                                     is.na(broadband_cost_month) & hh_size_recode==1 ~ complement_bband %>% filter(hh_size_recode==1) %>% select(mean_bband) %>% pull(),
                                                     is.na(broadband_cost_month) & hh_size_recode==2 ~ complement_bband %>% filter(hh_size_recode==2) %>% select(mean_bband) %>% pull(),
                                                     is.na(broadband_cost_month) & hh_size_recode==3 ~ complement_bband %>% filter(hh_size_recode==3) %>% select(mean_bband) %>% pull(),
                                                     is.na(broadband_cost_month) & hh_size_recode==4 ~ complement_bband %>% filter(hh_size_recode==4) %>% select(mean_bband) %>% pull(),
                                                     is.na(broadband_cost_month) & hh_size_recode==5 ~ complement_bband %>% filter(hh_size_recode==5) %>% select(mean_bband) %>% pull()
                                                     ))

```

# Other costs
```{r}
percentage_other_expenses <- 0.2

df_hlb_tic <- df_hlb_tic %>% mutate(other_month = (food_cost_month+housing_cost_month)*percentage_other_expenses ) 
  
```

#  ######################## 3 Data set  ########################

# Save data set of HLB with information of census tracts, income category and household combinations
# With this file, proceed to sample to create the synthetic population
```{r}

#save file in 
if(!file.exists( paste0("~/git/household_living_budget/documents/products/data_tables/", name_files, "_", state_name, "_hlb_no_tax_2022.csv") )){
  
  write.csv.gz(df_hlb_tic, paste0("~/git/household_living_budget/documents/products/data_tables/", name_files, "_", state_name, "_hlb_no_tax_2022.csv") , row.names = FALSE)
}

```


